The effects of binarity on planet occurrence rates measured by transit surveys

### 2018-02-14 paper/main.tex
improve text (minor)

\shortauthors{Bouma et al.}
\shorttitle{Binarity and Occurrence Rates}

@@

@@

\paragraph{The rate of Earth analogs} The immediately arresting result
is a ``bump'' in the apparent rate density at apparent radii below
$2r_\oplus$: the true rate for singles is less than the apparent rate.
If secondaries host as many planets as single stars, the occurrence of
these small planets is overestimated by $\approx 50\%$.  The magnitude
of the systematic error decreases if there are fewer planets around
secondaries; if secondaries host no planets at all, the apparent rate
is overestimated by only $\approx 10\%$.

@@

If the apparent radius of the system is above $2r_\oplus$, and planets
exist about primaries and secondaries at equal rates, then the planet
is $\approx\!3$ times more likely to orbit the primary than the
secondary.  If secondaries host half as many planets as primaries,
then a detected planet is $\approx\!5$ times more likely to orbit the
primary.

The situation for apparent radii below $2r_\oplus$ is more nuanced.
For the two cases in which secondaries host fewer planets than
primaries, detected planets in binaries are still always more likely
to orbit the primary.  However, if planets exist at the same rate in
primaries and secondaries, then there is a turn-over radius,
$\approx\!0.4r_\oplus$, below which more of the detected planets in
binaries come from secondaries~--~they have all been ``diluted down''
from larger true planetary radii!  Rephrased, going from apparent
radii of $2r_\oplus$~to~$1.4r_\oplus$, the fraction of detected
planets with binary companions increases by anywhere from
$6\%$~to~$12\%$, depending on the relative occurrence of planets about
primaries and secondaries.  We compare this prediction with recent
observations in Section~\ref{sec:discussion}.
%~\footnote{
%The increase can be derived analytically.  Let the fraction of total
%detected planets from primaries at a given apparent radius be
%$F_1(r\a)$, so $F_1(r\a) \equiv n_{\rm det}^1(r\a)/n_{\rm det}^{\rm
%tot}$.  Let the analogous quantity for secondaries be $F_2(r\a)$.  In
%the planet-less secondaries case ($F_2(r\a)=0$), one can analyze
%Eqs~\ref{eq:n_det_0} and~\ref{eq:n_det_1} semi-analytically and show
%that $F_1(2r_\oplus/\sqrt{2})/F_1(2r_\oplus) \approx 1.19$.
%check_companion_fractions.py, and LB's notes 2017/12/21.0 }.

@@

\paragraph{Hot Jupiter rate discrepancy}
While unresolved binaries may bias $\eta_\oplus$ measurements, our
approach suggests that they do not influence the hot Jupiter rate
discrepancy.  The ``discrepancy'' is that hot Jupiter occurrence rates
measured by transit surveys ($\approx 0.5\%$) are lower than those
found by radial velocity surveys ($\approx 1\%$; see
Table~\ref{tab:hj_rates}).

Though the disagreement is only weakly significant ($<3\sigma$), one
reason to expect a difference is that the RV sample is biased towards
metal-rich stars, which host more giant
planets~\citep{santos_spectroscopic_2004,fischer_planet-metallicity_2005,gould_frequency_2006}.
Investigating the discrepancy from the metallicty
angle,~\citet{guo_metallicity_2017} measured the {\it Kepler} field's
mean metallicity to be $[{\rm M/H}]_{\rm Kepler}= -0.045\pm0.009$,
which is lower than the California Planet Search's mean of $[{\rm
M/H}]_{\rm CPS}= -0.005\pm0.006$.  The former value agrees with that
found by \citet{dong_metallicities_2014}.  Refitting for the
metallicity exponent in $\Lambda_{\rm HJ} \propto 10^{\beta [{\rm
M/H}] }$, \citeauthor{guo_metallicity_2017}\! found $\beta = 2.1\pm
0.7$, and noted that the metallicity difference could account for
0.2\% of the needed 0.5\% difference between the measured CKS
and {\it Kepler}\ rates\footnote{ \citet{petigura_CKS_2017} recently
found $\beta = 3.4^{0.9}_{-0.8}$.  If true, this implies that
metallicity could account for about half of the ``hot Jupiter rate
discrepancy''.  }.  \citeauthor{guo_metallicity_2017}\!  concluded
that ``other factors, such as binary contamination and imperfect
stellar properties'' must also be at play.

@@

Occurrence rate calculations are beginning to incorporate the results
of high-resolution imaging surveys.  For example, recent studies have
reduced contamination in the ``numerator'' of the occurrence rate by
using~\citet{furlan_kepler_2017}'s catalog to test the effects of
removing KOI hosts with known companions
(\citealt{fulton_california-_2017}; \citealt{petigura_CKS_2017}).
This is a good first step, but the denominator remains uncorrected.

@@

Applying Equation~\ref{eq:general_Gamma_a}, we then showed that:
\begin{itemize}
%
\item If single, primary, and secondary stars all host the same number
  of planets per star, and the planet radius distribution is a power
    law with the exponent reported by~\citet{howard_planet_2012} at $r
    > 2r_\oplus$, then binarity influences apparent planet occurrence
    rates at the few percent level. This applies for apparent radii
    from $2r_\oplus$ to $17r_\oplus$ (Section~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with the
  \citet{howard_planet_2012} exponent above $2r_\oplus$ and a constant
    occurrence below $2r_\oplus$, there is a ``bump'' in the apparent
    rate density at $r\a < 2r_\oplus$. This can lead to
    overestimates in occurrence, $\Delta \Gamma_0 = |\Gamma_0 -
    \Gamma\a|/\Gamma_0$, of roughly $50\%$
    (Figure~\ref{fig:occ_rate_model_3_log}).  Although this is smaller
    than current systematic uncertainties on the occurrence rates of
    Earth-sized planets, this implies that binarity could eventually
    become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity ``fills in'' gaps in the radius distribution
  (Figure~\ref{fig:model_4}), to a degree that could affect precise
    measurements of the depth, width, and slope
    of~\citet{fulton_california-_2017}'s radius gap, in the event that
    planets are not carefully vetted with high-resolution imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates
  (Figure~\ref{fig:gaussian_HJ}).  This assumes a Gaussian planet
    radius distribution, similar to that reported
    by~\citet{petigura_CKS_2017}.
%
\item Detected planets with $r\a \gtrsim 0.5r_\oplus$ that are
  revealed by high-resolution imaging surveys to exist in binaries are
    more likely to orbit the primary (Figure~\ref{fig:frac_model_3}).
%
\item Near the ``break'' in occurrence rate as a function of planet
  radius ($\approx 2r_\oplus$), the fraction of detected planets with
    binary companions should increase by roughly $5$~-~$10\%$
    (Figure~\ref{fig:frac_model_3}).
\end{itemize}

>>>>

\shortauthors{Bouma et al.}
\shorttitle{Binary Biases}

@@

\item The contribution to the apparent rate density from
secondaries now depends on the mass of the secondary, $qM\a$. The
associated measure on the right-hand side of
Equation~\ref{eq:n2} becomes ${\rm d}(qM\a)$, and gives rise to an
extra multiplicative factor of $q$.

@@

\paragraph{The rate of Earth analogs} The immediately arresting result
is a ``bump'' in the apparent rate density at apparent radii below
$2r_\oplus$: the true rate for singles is less than the apparent rate.
If secondaries host as many planets as single stars, the occurrence of
these small planets is overestimated by about $50\%$.  The magnitude
of the systematic error decreases if there are fewer planets around
secondaries; if secondaries host no planets at all, the apparent rate
is overestimated by only $10\%$.

@@

If the apparent radius of the system is above $2r_\oplus$, and planets
exist about primaries and secondaries at equal rates, then the planet
is $3$ times more likely to orbit the primary than the
secondary.  If secondaries host half as many planets as primaries,
then a detected planet is $5$ times more likely to orbit the
primary.

The situation for apparent radii below $2r_\oplus$ is more nuanced.
For the two cases in which secondaries host fewer planets than
primaries, detected planets in binaries are still always more likely
to orbit the primary.  However, if planets exist at the same rate in
primaries and secondaries, then below apparent radii of $0.4r_\oplus$,
more of the detected planets in binaries come from secondaries~--~they
have all been diluted down from larger true planetary radii.
Rephrased, going from apparent radii of $2r_\oplus$~to~$1.4r_\oplus$,
the fraction of detected planets with binary companions increases by
anywhere from $6\%$~to~$12\%$, depending on the relative occurrence of
planets about primaries and secondaries.  We compare this prediction
with recent observations in Section~\ref{sec:discussion}.
%~\footnote{
%The increase can be derived analytically.  Let the fraction of total
%detected planets from primaries at a given apparent radius be
%$F_1(r\a)$, so $F_1(r\a) \equiv n_{\rm det}^1(r\a)/n_{\rm det}^{\rm
%tot}$.  Let the analogous quantity for secondaries be $F_2(r\a)$.  In
%the planet-less secondaries case ($F_2(r\a)=0$), one can analyze
%Eqs~\ref{eq:n_det_0} and~\ref{eq:n_det_1} semi-analytically and show
%that $F_1(2r_\oplus/\sqrt{2})/F_1(2r_\oplus) \approx 1.19$.
%check_companion_fractions.py, and LB's notes 2017/12/21.0 }.

@@

\paragraph{Hot Jupiter rate discrepancy}
While unresolved binaries may bias $\eta_\oplus$ measurements, our
approach suggests that they do not influence the hot Jupiter rate
discrepancy.  The ``discrepancy'' is that RV surveys find roughly
twice as many hot Jupiters per Sun-like star as do transit surveys
($\approx 1\%$~vs.~$\approx 0.5\%$, respectively).  We summarize hot
Jupiter rates reported by different surveys in
Table~\ref{tab:hj_rates}.

Though the disagreement is only weakly significant ($<3\sigma$), one
reason to expect a difference is that the RV sample is biased towards
metal-rich stars, which host more giant
planets~\citep{santos_spectroscopic_2004,fischer_planet-metallicity_2005,gould_frequency_2006}.
Investigating the discrepancy from the metallicty
angle,~\citet{guo_metallicity_2017} measured the {\it Kepler} field's
mean metallicity to be $[{\rm M/H}]_{\rm Kepler}= -0.045\pm0.009$,
which is lower than the California Planet Search's mean of $[{\rm
M/H}]_{\rm CPS}= -0.005\pm0.006$.  The former value agrees with that
found by \citet{dong_metallicities_2014}.  Refitting for the
metallicity exponent in $\Lambda_{\rm HJ} \propto 10^{\beta [{\rm
M/H}] }$, \citeauthor{guo_metallicity_2017}\! found $\beta = 2.1\pm
0.7$, and noted that the metallicity difference could account for
0.2\% of the needed 0.5\% difference between the measured CKS
and {\it Kepler}\ rates\footnote{ \citet{petigura_CKS_2017} recently
found $\beta = 3.4^{0.9}_{-0.8}$.  If true, this implies that
metallicity could account for about half of the hot Jupiter rate
discrepancy.  }.  \citeauthor{guo_metallicity_2017}\!  concluded
that ``other factors, such as binary contamination and imperfect
stellar properties'' must also be at play.

@@

Occurrence rate calculations are beginning to incorporate the results
of high-resolution imaging surveys.  For example, recent studies have
reduced contamination in the numerator of the occurrence rate by
using~\citet{furlan_kepler_2017}'s catalog to test the effects of
removing KOI hosts with known companions
(\citealt{fulton_california-_2017}; \citealt{petigura_CKS_2017}).
This is a good first step, but the denominator remains uncorrected.

@@

Applying Equation~\ref{eq:general_Gamma_a}, we then showed that:
\begin{itemize}
%
\item If single, primary, and secondary stars all host the same number
  of planets per star, and the planet radius distribution is a power
    law with the exponent reported by~\citet{howard_planet_2012} at $r
    > 2r_\oplus$, then binarity influences apparent planet occurrence
    rates at the few percent level. This applies for apparent radii
    from $2r_\oplus$ to $17r_\oplus$ (Section~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with the
  \citet{howard_planet_2012} exponent above $2r_\oplus$ and a constant
    occurrence below $2r_\oplus$, the rate density for radii below
    $2r_\oplus$ is overestimated. The overestimate, $\Delta \Gamma_0 =
    |\Gamma_0 - \Gamma\a|/\Gamma_0$, can easily reach $50\%$
    (Figure~\ref{fig:occ_rate_model_3_log}).  Although this is smaller
    than current systematic uncertainties on the occurrence rates of
    Earth-sized planets, this implies that binarity could eventually
    become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity blurs sharp features in the radius distribution
  (Figure~\ref{fig:model_4}), to a degree that could affect precise
    measurements of the depth, width, and slope of
    the~\citet{fulton_california-_2017} radius gap.
    %in the event that
    %planets are not carefully vetted with high-resolution imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates
  (Figure~\ref{fig:gaussian_HJ}).  This assumes a Gaussian planet
    radius distribution, similar to that reported
    by~\citet{petigura_CKS_2017}.
%
\item Detected planets with $r\a \gtrsim 0.5r_\oplus$ that are
  revealed by high-resolution imaging surveys to exist in binaries are
    more likely to orbit the primary (Figure~\ref{fig:frac_model_3}).
%
\item Near the break in occurrence rate as a function of planet
  radius ($\approx 2r_\oplus$), the fraction of detected planets with
    binary companions should increase by roughly $5$~-~$10\%$
    (Figure~\ref{fig:frac_model_3}).
\end{itemize}
